### 2020-04-12 TOG/note.tex
simplify

\section{Blog}

\begin{description}
\item[December 18, 2016]
{
\qisun{
I am starting to double this as a tutorial on how to use paper drafts to manage project progress.
This is still work in progress and not yet done, but I hope to finish a draft within the next few months.
}%qisun
}%item

\end{description}