### 2020-12-31 TOG/title_author.tex

\title{Using Paper Drafts to Manage Research Projects}

\author{Anonymous Authors}