### 2021-01-05 TOG/title_author.tex

\title{Representing Scenes as Light Fields for Real Time Rendering in Virtual Reality Display}
\title{Immersive 6DoF Streaming with Gaze-Contingent Neural Synthesis} 
\title{Gaze-Contingent Neural Synthesis for Low Latency VR in-the-Cloud} 
% Representing Scenes as Neural Radiance Fields for View Synthesis

% perceptually-aware neural synthesis for low latency 6dof VR?
% perceptual neural rendering for ultra fast 6dof VR

\author{Anonymous Authors}