### 2021-01-12 TOG/1-introduction.tex

\section{Introduction}

\note{
What problem we are trying to solve.
Why it is important, and why people should care.
}%note

With the advent of remote desktops and streaming services, cloud-based computing has shown unprecedented potential reshaping traditional computer graphics pipeline by completely relieving end-user's computational workload, local storage volume, thus advance device portability. Recent years has seen growing interactive streaming services such as Google Stadia or NVIDIA Shield. 
However, these off-the-shelf services have only been deployed in less latency-sensitive platforms such as TV-based game consoles, due to the inevitable latency (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}. Immersive media such as virtual and augmented reality (VR/AR), however, demand remarkably lower latency and higher fidelity/speed for sickness-free and natural experience. 
Due to the consequently massive spatial-temporal high-quality data volume, there is no available solution today that is capable of streaming such dense content to the individual end user.

\note{
What prior works have done, and why they are not adequate.
(Note: this is just high level big ideas. Details should go to a previous work section.)
}%note
\qisun{(Jan 5, 2021) We should include two main fields: neural representation and foveated rendering}

Recent years have seen a surge in neural representations as an alternative to compact 3D representations such as point clouds and polygonal meshes, enabling compact assets. Application of neural representations include voxelization \cite{sitzmann2019deepvoxels}, distance fields \cite{sitzmann2019metasdf,mildenhall2020nerf} and many more. However, existing solutions suffer from either high time consumption or low image fidelity.

To accelerate real-time rendering tailored for VR, gaze-contingent foveated rendering approaches have shown remarkable effectiveness of presenting imagery of perceptual quality identical to full resolution rendering~\cite{Patney:2016:TFR}. This is achieved by harnessing high FoV head-mounted displays and high precision eye-tracking technologies. Unfortunately, foveated rendering approaches still require full local 3D asset storage, resulting in a significant loading time and memory shortage.

\note{
What our method has to offer, sales pitch for concrete benefits, not technical details.
Imagine we are doing a TV advertisement here.
}%note

In this paper, we present the first gaze-contingent real-time 6DoF neural synthesis of scenes, aiming at the next generation cloud-based wearable immersive platforms.
Specifically, our method significantly reduces 3D data storage (\warning{xxx ms}), and rendering time (\warning{xxx ms}) compared with the existing methods. It enables low-latency interactive VR/AR 6DoF viewable 3D virtual content without compromising perceived quality. Tailoring for future light weight, power and memory efficient wearable display systems, expected to be active all-day long akin to today's everyday-use mobile phones, our method achieves minimal local storage for any high quality 3D scenes.

\note{
Our main idea, giving people a take home message and (if possible) see how clever we are.
}%note

In this work, we take an interdisciplinary approach achieving 6-DoF deep neural representation of complex scenes and gaze-contingent real-time rendering. This is realized by our novel data representation and a coarse-to-fine prediction of perceptually high quality imagery while balancing the latency.

\note{
Our algorithms and methods to show technical contributions and that our solutions are not trivial.
}%note

Specifically, we represent an arbitrary virtual 3D scene as a compact neural representation in a viewer centered inside-out fashion. 
We account for the challenges such as low quality/resolution/performance common to existing 3D deep neural representation approaches, we further leverage the human visual perception to synthesize gaze-contingent and vision-matched immersive frames, without compromising the speed and image fidelity. 
The method is further optimized with in the spatio-temporal domain to balance the latency and precision.

\note{
Results, applications, and extra benefits.
}%note

We validate our approach by conducting both numerical analysis and user evaluations on commercially available AR/VR display devices. They both reveal the effectiveness of our method on delivering a real-time VR viewing experience, perceptually identical to an arbitrarily high quality rendered 3D scene, without compromising on local storage volume. Our main contributions include the following:
\begin{itemize}
    \item A low-latency and significantly low-storage 6DoF immersive application that preserves perceptual quality
    \item A gaze-contingent neural representation and synthesis method tailored for first-person viewing
    \item A spatio-temporally joint optimization for balancing latency and perceptual quality
\end{itemize}