### 2021-01-12 TOG/2-prior.tex

\section{Related Work}
\label{sec:prior}

\subsection{Deep View Synthesis}

\subsection{Gaze-Contingent Rendering}
\cite{Guenter:2012:F3G,Patney:2016:TFR,Sun:2017:PGF,Kaplanyan:2019:DNR}