### 2021-01-21 TOG/fig_lpips.tex

\begin{figure*}
    \centering
    \subfloat[gas]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips-fov-gas-group.pdf}}\label{lpips:gas}
    \subfloat[minecraft]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenemc.pdf}}

    \subfloat[studyroom]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenestudyroom.pdf}}
    \subfloat[bedroom]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenebedroom.pdf}}
    \caption{LPIPS analysis of all scenes as well as comparison of our method, NeRF, and foveated GT on each eccentricity. $^*$: $p<.05$, $^{**}$: $p<.01$.}
    {We found a significant difference in \warning{ecc range} between \warning{stimulus} and \warning{stimulus};}
    \label{fig:lpips}
\end{figure*}