### 2021-01-28 TOG/ccs.tex

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010147.10010371</concept_id>
<concept_desc>Computing methodologies~Computer graphics</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003120.10003121.10003124.10010866</concept_id>
<concept_desc>Human-centered computing~Virtual reality</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept_id>10010147.10010371.10010387.10010393</concept_id>
<concept_desc>Computing methodologies~Perception</concept_desc>
<concept_significance>500</concept_significance>
</concept>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Perception}
\ccsdesc[500]{Human-centered computing~Virtual reality}

\keywords{virtual reality; gaze-contingent graphics; neural representation}