### 2021-01-28 TOG/fig_ray_bounding.tex

\begin{figure}
    \centering
    \subfloat[w/o bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_without_bound.png}}\hspace{1em}
    \subfloat[w bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_with_bound.png}}
    \Caption{Comparison of ray marching bounding.}
    {%
    }
    \label{fig:method:bounding}
\end{figure}