### 2021-01-28 TOG/teaser_fig.tex

%\begin{figure}[htb]
\begin{teaserfigure}
  \centering
  \subfloat[scene representation]{
    \label{fig:teaser:scene}
    % \includegraphics[height=4.5cm]{TOG/figs/scene.pdf}
    \includegraphics[trim=100 0 120 0,clip,height=5.5cm]{TOG/figs/teaser/teaser_1a_wc.pdf}
  }%subfloat
  \subfloat[overall quality]{
    \label{fig:teaser:latency}
    % \includegraphics[height=4.5cm]{TOG/figs/teaser_1b_update.pdf}
    % text version
    % \includegraphics[height=5.5cm]{TOG/figs/teaser/overall_crop_zh_large.pdf}
    % color version
    \includegraphics[height=5.5cm]{TOG/figs/teaser/overall_crop_color_s.pdf}
    % nerf: 15+s/per infer
  }%subfloat
  \subfloat[foveal quality]{
    \label{fig:teaser:quality}
    % \includegraphics[height=4.5cm]{TOG/figs/teaser_1c.pdf}
    % text version
    % \includegraphics[height=5.5cm]{TOG/figs/teaser/overall_flat_zh_large.pdf}
    % color version
    \includegraphics[height=5.5cm]{TOG/figs/teaser/overall_flat_color_s.pdf}
  }%subfloat
 \Caption{Illustration of our gaze-contingent neural synthesis method.}
 {% 
\subref{fig:teaser:scene} Visualization of our active-viewing-tailored egocentric neural scene representation.
\subref{fig:teaser:latency} Our method allows for extremely low data storage of high quality 3D scene assets and achieves high perceptual image quality and low latency for interactive virtual reality. Here, we compare our method with the state-of-the-art alternative neural synthesis \protect\cite{mildenhall2020nerf} and foveated rendering solutions \cite{perry2002gaze,Patney:2016:TFR} and demonstrate that our method significantly reduces time consumption and data storage (more than $99\%$ reduction for both) for first-person immersive viewing. The orange circle indicates viewer's gaze position. We also show zoomed-in views of the foveal images generated by various methods in \subref{fig:teaser:quality} and show that our method is superior compared to existing methods in achieving high perceptual image fidelity.
%\qisun{(Jan 24, 2021) Let's add some latency numbers to
%\zh{15s/infer for NeRF}\subref{fig:teaser:latency}?}
%\qisun{(Jan 25) Can I get your help on adding the numbers with the fancy font to the figures directly in \subref{fig:teaser:latency}?}
 }
 \label{fig:teaser}
%\end{figure}
\end{teaserfigure}