### 2021-03-09 TOG/fig_blending.tex

\begin{figure*}[ht]
    \centering
    \subfloat[network overlay]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/overlap2.pdf}\label{fig:method:blending:overlay}}\hspace{1em}
    \subfloat[blending]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended2.pdf}\label{fig:method:blending:blending}}\hspace{1em}
    \subfloat[contrast enhancement]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended_enhanced2.pdf}\label{fig:method:blending:contrast}}
    \Caption{Our multi-layer real-time rendering system.}
    {%
    \subref{fig:method:blending:overlay} shows the original 3-layer image output overlaid.
    \subref{fig:method:blending:blending} shows the result with cross-layer blending that enables smooth appearance on the edges.
    \subref{fig:method:blending:contrast} shows the contrast-enhanced rendering pass in the periphery.
    }
    \label{fig:method:blending}
\end{figure*}