### 2021-03-09 TOG/fig_notations.tex

\begin{figure}
    \centering
    \includegraphics[width=0.96\linewidth]{TOG/figs/parameters.pdf}
    % \subfloat[Notations]{\includegraphics[width=0.96\linewidth]{TOG/figs/parameters.pdf}\label{fig:variables}}

    % \subfloat[Correlation between loss and parameters]{\includegraphics[width = 0.96\linewidth]{example-image-a}}
    \Caption{Coordinate system and variable annotations.}
    {%
    We partially visualize our 3D full spherical representation with an example of $\sphereNum=3$ spheres. Variables and equations are annotated at corresponding positions.
    }
    \label{fig:notations}
\end{figure}