### 2021-03-09 TOG/fig_optimization.tex

\begin{figure}[t]
    \centering
    \subfloat[quality]{\includegraphics[width=0.48\linewidth]{TOG/figs/latency_vs_quality/heat_quality2.pdf}\label{fig:optimization:heat:quality}}
    \subfloat[latency]{\includegraphics[width=0.48\linewidth]{TOG/figs/latency_vs_quality/heat_time2.pdf}\label{fig:optimization:heat:time}}

    \subfloat[quality-only]{\includegraphics[width=0.3\linewidth]{TOG/figs/latency_vs_quality/quality_only.pdf}\label{fig:optimization:quality}}\hspace{0.5em}
    \subfloat[latency-only]{\includegraphics[width=0.3\linewidth]{TOG/figs/latency_vs_quality/latency_only.pdf}\label{fig:optimization:time}}\hspace{0.5em}
    \subfloat[ours]{\includegraphics[width=0.3\linewidth]{TOG/figs/latency_vs_quality/our.pdf}\label{fig:optimization:our}}
    \Caption{Latency-quality joint-optimization.}
    {%
    \protect\subref{fig:optimization:heat:quality}/\protect\subref{fig:optimization:heat:time} plots $\imgSpaceError$ in \protect\Cref{eq:imageError}/latency of an example foveal network during the optimization process. (in milliseconds\nothing{ under pytorch implementation}) .
    The values are computed with various settings of $\mlpChannelNum\times\mlpLayerNum$ (X-axis) and $\sphereNum$ (Y-axis). The second row indicates a corresponding foveal image under different settings. Our method balances both perceptual quality and latency.\qisun{note to myself: add deeper discussion and insights here.}
    }
    \label{fig:optimization}
\end{figure}