### 2021-03-09 TOG/fig_ray_bounding.tex

\begin{figure}[hb]
    \centering
    \subfloat[infinite upper bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_without_bound.png}\label{fig:impl:bounding:off}}\hspace{1em}
    \subfloat[limited bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_with_bound.png}\label{fig:impl:bounding:on}}
    \Caption{The effect of ray marching bounding.}
    {%
    Compared to infinite ray marching \subref{fig:impl:bounding:off}, bounding the ray marching further improves visual quality in \subref{fig:impl:bounding:on}.
    }
    \label{fig:impl:bounding}
\end{figure}