### 2021-03-09 TOG/note.tex

\section{Blog}

% \item[Jan 28, 2016]
{
\qisun{
\begin{itemize}
    \item dataset's diversity and generality
    \item algorithm overview
    \item foveal vs periphery in optimization

\end{itemize}
}%qisun
}%item

\begin{description}
\item[December 18, 2016]
{
\qisun{
I am starting to double this as a tutorial on how to use paper drafts to manage project progress.
This is still work in progress and not yet done, but I hope to finish a draft within the next few months.
}%qisun
}%item

\end{description}